---
title: "Introduction and basic workflow for EflowStats"
author: "T. Joe Mills"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{EflowStats Intro}
  \usepackage[utf8]{inputenc}
---
  
  
  
## Overview
  
`EflowStats` is a reimplementation of the Hydrologic Index Tool (HIT; Henriksen et al, 2006) for calculating 171 hydrologic indices for stream classification analysis. Additionally, `EflowStats` can calculate 7 additional statistics used for streamflow classification referred to as the "Magnificent Seven" (MAG7, Archfield et al., 2013). Unlike the original HIT, EflowStats has been redesigned to ingest general hydrologic timeseries data and is not restricted to data formats used by the USGS National Water Information System.
  
##A basic workflow 
  
`EflowStats` can be used with the `dataRetrieval` R package to seamlessly calculate HIT and MAG7 statistics for any USGS stream gage. 
  
  
```{r message=FALSE, echo=TRUE, eval=FALSE}
library(dataRetrieval)
library(EflowStats)

#Get some data
dailyQ <- readNWISdv(siteNumber = "04085427",
                     parameterCd = "00060",
                     startDate = "2000-10-01",
                     endDate = "2012-9-30")

#Get drainage area
siteInfo <- readNWISsite(siteNumber = "04085427")
drainageArea <- siteInfo$drain_area_va

#Get peak flows
peakFlows <- readNWISpeak(siteNumber = "04085427",
                          startDate = "2000-10-01",
                          endDate = "2012-9-30")

```

Users with their own time series data can use the package as long as the data can be coerced into a `data.frame` with dates of class _Date_ in the first column, and stream flow values of class _Numeric_ in the second. The code below shows steps to calculate all hydrologic indices available from the package.

```{r message=FALSE, echo=TRUE, eval=FALSE}
#Check data for completeness
dailyQClean <- validate_data(dailyQ[c("Date","X_00060_00003")],
                             yearType="water")

#Get flood recurence threshold
floodThresh <- get_peakThreshold(dailyQClean[c("date","discharge")],
                                 peakFlows[c("peak_dt","peak_va")])

#Calculate all hit stats
calc_allHITOut <- calc_allHIT(dailyQClean,
                              drainArea=drainageArea,
                              floodThreshold=floodThresh)

#Calculate mag7 stats
magnifStatsOut <- calc_magnifSeven(dailyQClean,yearType="water",digits=3)


```


Definitions of the statistics calculated by EflowStats can be found in Henriksen et al, 2006 and Archfield et al., 2013 as well as in the package function documentation. Users who have used the hydroecological integrity assessment process software or previous versions of EflowStats (<v1.0) may find slight differences in the output of the new package (>v1.0). The package has undergone extensive review and every effort has been made to ensure faithful reproduction of the statistics as defined in Henriksen et al, 2006. Users interested in records of correspondence between the old package (<v1.0) and the new package (>v1.0) are directed to the summary in the [Package Discrepancies vignette](packageDiscrepencies.html). Additionally, a comparison with output from the old package and the new package can be found at  [this issue](https://github.com/USGS-R/EflowStats/issues/132) and the history of review documented in [other closed issues.](https://github.com/USGS-R/EflowStats/issues?q=is%3Aissue+is%3Aclosed).


##References
Henriksen, J. A., Heasley, J., Kennen, J.G., and Niewsand, S. 2006, Users’ manual for the hydroecological
integrity assessment process software (including the New Jersey Assessment Tools): U.S. Geological
Survey, Biological Resources Discipline, Open File Report 2006-1093, 71 p. 

Archfield, S.A., J.G. Kennen, D.M. Carlisle, and D.M. Wolock. 2013. An Objective and Parsimonious Approach for Classifying Natural Flow Regimes at a Continental Scale. River Res. Applic. doi: 10.1002/rra.2710
